Quality Definitions of Qualities

### 2019-10-09 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
my 4 definitions.

\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{amsmath}

@@

@@

To assess the current state of software development, and to understand how
future changes impact software development, we need a clear definition of what
we mean by quality.  The concept of quality is decomposed into a set of separate
qualities.  This set of qualities can be applied to the software product, the
software artifacts (documentation, test cases, etc) and to the software
development process itself.

@@

Usability can be a problem.  Different users, solving the same physical problem,
using the same software, can come up with different answers, due to differences
in parameter selection~\cite[p.~370]{Roache1998}.  To reduce misuse, a rational
process must state expected user characteristics, modelling assumptions,
definitions and the range of applicability of the code.

@@

Understandability is necessary, since reviewers can only certify something they
understand.  Scientific software developers have the view ``that the science a
developer embeds in the code must be apparent to another scientist, even ten
years later''~\cite{Kelly2013}.  Understandability applies to the documentation
and code, while usability refers to the executable software.  Documentation that
follows a rational process is the easiest to follow.

@@

The documentation should be developed in such a way that it is easily modifiable
so that likely future changes do not destroy the structure of the document. Also
it should be easy to reflect the change, wherever needed in the document to
maintain consistency, traceability and completeness. For documentation to be
modifiable, its format must be structured in a way that repetition is avoided
and cross-referencing is employed.

@@

Documentation should be traceable, as this facilitates maintenance and
review. If a change is made to the design or code of the software, then all the
documentation relating to those segments have to be modified.  This property is
also important for recertification.

>>>>

\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}

@@

\theoremstyle{definition}
\newtheorem{defn}{Definition}

@@

To assess the current state of software development, and to understand how
future changes impact software development, we need a clear definition of what
we mean by quality.  The concept of quality is decomposed into a set of separate
components that together make up ``quality''. Unfortunately, these are
called \emph{qualities}. These are associated to the software product,
the software artifacts (documentation, test cases, etc) and to the software
development process itself, and combinations thereof.

@@

ISO defines usability as
\begin{quote}
The extent to which a product can be used by specified users to achieve
specified goals with effectiveness, efficiency, and satisfaction in a specified
context of use.
\end{quote}

Nielsen and (separately) Schneidermann have defined usability as part of usefulness and
is composed of
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the
  first time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the economics
sense of the word).

There are two ISO standards covering this, namely ISO/TR 16982:202 and ISO 9241.

The Interaction Design Foundation~\url{https://www.interaction-design.org/literature/topics/usability}
further lists the following desirable outcomes:
\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in using
the user interface during the first contact with the website. For example, if a
travel agent’s website is a well-designed one, the user should be able to move
through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
website. If a user has the goal of booking a flight, a good design will guide
him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
subsequent visits. So, a good design on the travel agent’s site means the user
should learn from the first time and book a second ticket just as easily.
\end{enumerate}

One core reference, for definitions and metrics, is Bevan~\cite{bevan1995measuring}.

@@

Understandability is artifact-dependent. What it means for a user-interface (graphical
or otherwise) to be understandable is wildly different than what it means for the code,
and even the user documentation.

The literature here is thin and scattered.  More work will need to be done to find
something useful.

Interestingly, the business literature seems to have taken more care to define this.
Here we encounter
\begin{quote}
Understandability is the concept that X should be presented
so that a reader can easily comprehend it.
\end{quote}
At least this brings in the idea that the \emph{reader} is actively involved, and
indirectly that the reader's knowledge may be relevant, as well as the
``clarity of exposition'' of X.

Section 11.2 of~\cite{adams2015nonfunctional} does have a full definition.

@@

Here we do seem to have a simple, if somewhat uninformative, definition:

\begin{defn}
Modifiability is the degree of ease at which changes can be made to a system,
and the flexibility with which the system adapts to such changes.
\end{defn}

IEEE Standard 610 seems to speak about this. (which is superceded?)

@@

Here the Wikipedia page~\url{https://en.wikipedia.org/wiki/Traceability} is actually
rather informative, especially as it also lists how this concept is used in other domains.
A generic definition that is still quite useful is
\begin{defn}
The capability (and implementation) of keeping track of a given set or type of
information to a given degree, or the ability to chronologically interrelate
uniquely identifiable entities in a way that is verifiable.
\end{defn}
By specializing the above to software artifacts, ``interrelate'' to 
``why is this here'' (for forward tracing from requirements), this does
indeed give what is meant in SE.

Various standards (DO178C, ISO 26262, and IEC61508) explicitly mention it.

24765-2017 - ISO/IEC/IEEE International Standard - Systems and software engineering--Vocabulary
has a full definition, namely
\begin{enumerate}
\item the degree to which a relationship can be established between two or more
products of the development process, especially products having a
predecessor-successor or master-subordinate relationship to one another;
\item
the identification and documentation of derivation paths (upward) and
allocation or flowdown paths (downward) of work products in the work product
hierarchy;
\item the degree to which each element in a software development
product establishes its reason for existing; and discernible association
among two or more logical entities, such as requirements, system elements,
verifications, or tasks.
\end{enumerate}
